Reanalysis Step 5: Outcome Analysis with Three-Way Interaction

### 2025-09-01 reanalysis-5-outcome_analysis-interaction.qmd
debug ix

## Pooling Methodology for Survey-Weighted Multiple Imputation

This analysis combines:
1. **Multiple Imputation** to handle missing data
2. **IPTW weighting** from propensity scores to handle confounding
3. **Survey methods** to handle clustering (multiple eyes per patient)

### Rubin's Rules for Combining Results

Following Rubin (1987) and Barnard & Rubin (1999), for each parameter θ, we combine estimates across m imputations:

1. **Point estimate**: $\bar{Q} = \frac{1}{m}\sum_{i=1}^{m} \hat{\theta}_i$

2. **Within-imputation variance**: $\bar{U} = \frac{1}{m}\sum_{i=1}^{m} \hat{SE}_i^2$

3. **Between-imputation variance**: $B = \frac{1}{m-1}\sum_{i=1}^{m} (\hat{\theta}_i - \bar{Q})^2$

4. **Total variance**: $T = \bar{U} + (1 + \frac{1}{m})B$

5. **Degrees of freedom** (Barnard-Rubin adjustment for finite samples):
   - $r = (1 + \frac{1}{m}) \frac{B}{\bar{U}}$ (relative increase in variance)
   - $df_{old} = (m-1) \times (1 + \frac{1}{r})^2$ (large sample df)
   - $\lambda = \frac{r + 2/(df_{com} + 3)}{r + 1}$ (fraction of missing information)
   - $df_{obs} = \frac{df_{com} + 1}{df_{com} + 3} \times df_{com} \times (1 - \lambda)$ (observed df)
   - $df_{adj} = \frac{df_{old} \times df_{obs}}{df_{old} + df_{obs}}$ (adjusted df)

Where $df_{com}$ is the complete-data degrees of freedom.

**References:**
- Rubin, D.B. (1987). *Multiple Imputation for Nonresponse in Surveys*. Wiley.
- Barnard, J. & Rubin, D.B. (1999). Small-sample degrees of freedom with multiple imputation. *Biometrika*, 86(4), 948-955.

### Implementation Using mitools

The `mitools` package (Lumley, 2020) provides robust implementations of Rubin's rules and extensions for survey data:

- **`MIcombine()`**: Implements standard Rubin's rules with Barnard-Rubin adjustment
- **`with.imputationList()`**: Applies analyses across imputations
- **`pool.scalar()`**: Combines scalar quantities (e.g., R²)
- **`MIextract()`**: Extracts results from fitted models

This analysis uses `mitools` rather than manual implementation to ensure accuracy and handle edge cases properly.

### Special Considerations for Survey Weights

- Standard errors from `svyglm` already account for IPTW weights
- Pooling preserves the design-based variance estimation
- Joint tests use Li-Raghunathan-Rubin method for χ² statistics (Li et al., 1991)

@@

# Pivot to wide format for better readability with percentages
library(tidyr)
crosstab_wide <- crosstab_df |>
  pivot_wider(names_from = Lapse_Status, 
              values_from = Count,
              values_fill = 0) |>
  arrange(DR_Severity, Treatment) |>
  mutate(
    Total = `0` + `1`,
    No_Lapse_Pct = paste0(round(100 * `0` / Total, 1), "%"),
    Lapsed_Pct = paste0(round(100 * `1` / Total, 1), "%"),
    Row_Pct = paste0(round(100 * Total / total_n, 1), "%")
  ) |>
  select(DR_Severity, Treatment, 
         No_Lapse = `0`, No_Lapse_Pct,
         Lapsed = `1`, Lapsed_Pct,
         Total, Row_Pct)

>>>>

## Analysis Overview

This analysis examines the three-way interaction between:
- **Lapse status** (primary exposure)
- **DR severity** (effect modifier)
- **Treatment status** (effect modifier)

The model uses:
- Multiple imputation to handle missing data
- IPTW weights from propensity scores
- Survey methods for clustered data (multiple eyes per patient)
- Pooling via Rubin's rules across imputations

@@

# Pivot to wide format for better readability with percentages
library(tidyr)
crosstab_wide <- crosstab_df |>
  pivot_wider(names_from = Lapse_Status, 
              values_from = Count,
              values_fill = 0,
              names_prefix = "Lapse_") |>  # Add prefix to avoid numeric column names
  arrange(DR_Severity, Treatment) |>
  mutate(
    Total = Lapse_0 + Lapse_1,
    No_Lapse_Pct = paste0(round(100 * Lapse_0 / Total, 1), "%"),
    Lapsed_Pct = paste0(round(100 * Lapse_1 / Total, 1), "%"),
    Row_Pct = paste0(round(100 * Total / total_n, 1), "%")
  ) |>
  select(DR_Severity, Treatment, 
         No_Lapse = Lapse_0, No_Lapse_Pct,
         Lapsed = Lapse_1, Lapsed_Pct,
         Total, Row_Pct)
